---
title: "R Notebook"
output: html_notebook
---

# Introduction

This document was created using R-Notebook, which aids in creating 
descriptive, reproducible analyses.

## The tidyverse

The packages `dplyr` and `tidyr` are used throughout this analysis. These packages
are from the `tidyverse`, which uses a slightly different syntax than base R. 
The `tidyverse` syntax is intended to reduce typing and to produce more readable
code. Although the syntax takes some time to learn, I find these packages
to be very helpful for manipulating data. Whenever I use these functions in this
document I have tried specify the package that they come from by adding the prefix
`dplyr::` or `tidyr::`. The prefix is not necessary if you load the package using
`library()` but I thought it might help to clarify which package the function
is derived from. 

The `tidyverse` syntax is referred to as Non-Standard Evaluation (NSE). NSE 
allows the user to refer to columns by name without specifying the
data frame for each column and without having to quote the column name.
For example, imagine you want to keep all rows from your data frame,
`df`, where `col.x == TRUE`. In base R, you could write `df[df$col.x == TRUE, ]`
to filter your data frame. The `dplyr` function, `filter`, is more descriptive 
of the action you want to perform and only requires the user to specify the
data frame once. To perform the same filter as above with `dplyr`, 
you would write `filter(df, col.x == TRUE)`. Another useful function from 
`dplyr` is `select`. Imagine `df` has 26 columns and you only want to keep
columns `col.x`, `col.y`, and `col.z`. In base R, you could write 
`df[, c("col.x", "col.y", "col.z")]`. To perform the same action with `dlyr` 
you would use `select(df, col.x, col.y, col.z)`. The `dplyr` function are 
more descriptive and generally require less typing.

Another useful feature of the `tidyverse` is the `magrittr` pipe function, `%>%`. 
The pipe allows the user to chain multiple functions together. The example provided
by Wickham and Grolemund (2016) is `x %>% f(y)` is equivalent to `f(x, y)`. 
Wickham and Grolemund (2016) also suggest reading `%>%` as "then". The following
example would be read as specify `df` as your data frame, then select columns 
`col.x`, `col.y`, and `col.z`, and then filter rows so that `col.x == TRUE`.

```{r eval=FALSE}
df %>% 
select(col.x, col.y, col.z) %>% 
filter(col.x == TRUE)
```

# Preparation

## Import Packages
Load `dplyr` and `tidyr` for data manipulation.
```{r, message=FALSE}
library(tidyverse)
library(tidyr)
library(rprojroot)
```

The `Benthos` package is the package I am currently developing. 
`Benthos` is stored on github and can be installed by using 
`devtools::install_github("zsmith27/Benthos", ref = "nse")`. 
If you receive an error when attempting to install from GitHub, 
try running `install.packages(c("devtools","curl", "httr"))` and
then `devtools::install_github("zsmith27/Benthos", ref = "nse")` again. 
If the issue persists please inform me on my GitHub page.

```{r}
library(mmir)
```