---
title: "R Notebook"
output: html_notebook
---

# Metric Calculation

Metrics are calculated using the `Benthos` package that I am developing.

## Sequence Metrics

The `Benthos` function `seq_taxa` allows the user to sequence across a 
specified vector of column names and calculate the percentage, abundance, 
or richness of each taxon in the specified columns. The vector of column 
names currently being used is `hier.vec`. You will notice that this vector 
does not just include taxonomic columns, but also, taxonomic attribute columns. 
Each taxonomic attribute is treated like a unique taxon, and therefore, the 
percentage, abundance, and richness of each taxonomic attribute can also be 
calculated. The `seq_taxa` function can only do one `job` at a time. The `dplyr` 
function `bind_cols` is used to append columns together from each `seq_taxa` call. 
Each metric calculated with the `seq_taxa` function will be named with a prefix 
designated to the specified `job` ("pct" = percentage, "abund" = abundance, and 
"rich" = richness) followed by the taxonomic name or taxonomic attribute name.

```{r, message=FALSE}
hier.vec <- c("water_type", "tolerance", "native", "trophic_class",
              "class", "superorder", "order", "suborder", "superfamily",
              "family", "subfamily", "tribe", "genus", "species")

seq.df <- Benthos::seq_taxa(fish.long, site, count, hier.vec, job = "pct") %>% 
  dplyr::bind_cols(Benthos::seq_taxa(fish.long, site, count, hier.vec, job = "abund")) %>% 
  dplyr::bind_cols(Benthos::seq_taxa(fish.long, site, count, hier.vec[!hier.vec %in% "species"],
                     high.taxa.col = "species", job = "rich"))
```

## Additional Metrics

Create a new data frame to store metric values. Keep only the columns 
that will be useful in subsequent analyses.
```{r}
metrics.df <- fish.long %>% 
  select(site, gradient, basin, drainage_size) %>% 
  distinct()
```

The `seq_taxa` function cannot be used to calculate all metrics. Below 
total species richness, weighted species richness, total abundance, and 
the percent of dominant taxa (1, 2, 3, 4, and 5). `pct_`, `abund_`, and 
`rich_` `insectivores` was calculated as the sum of the corresponding 
`benthic_insectivore`, `water_column_insectivore`, and `pi` metrics calculated 
with `seq_taxa`.
```{r}
#==============================================================================
metrics.df <- dplyr::bind_cols(metrics.df, seq.df)

metrics.df <- metrics.df %>% 
  mutate(richness = taxa_rich(fish.long, site, count, family, species)) %>% 
  left_join(env.df[, c("site", "width")], by = "site") %>% 
#  mutate(weighted_richness =
#           case_when(
#             width <= 4 ~ richness + 2,
#             width > 4 & width <= 9 ~ as.double(richness),
#             width > 9 & width <= 19 ~ richness - 2,
#             width > 19 ~ richness - 4,
#             is.na(width) ~ as.numeric(NA),
#             TRUE ~ as.numeric(NA)
#           )) %>% 
  select(-width) %>% 
  mutate(abundance = taxa_abund(fish.long, site, count, species),
         shannon = taxa_div(fish.long, site, count, species, job = "shannon", base = 2),
         simpson = taxa_div(fish.long, site, count, species, job = "simpson"),
         hill_2 = taxa_div(fish.long, site, count, species, job = "hill", q = 2),
         pct_dom1 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 1),
         pct_dom2 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 2),
         pct_dom3 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 3),
         pct_dom4 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 4),
         pct_dom5 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 5),
         pct_insectivores = sum(pct_benthic_insectivore,
                                pct_water_column_insectivore,
                                pct_pi),
         abund_insectivores = sum(abund_benthic_insectivore,
                                abund_water_column_insectivore,
                                abund_pi),
         rich_insectivores = sum(rich_benthic_insectivore,
                                rich_water_column_insectivore,
                                rich_pi))
```


`r ncol(metrics.df) - 4` metrics were calculated.