---
title: "R Notebook"
output: html_notebook
---

## Station Data

### Import

The station and environmental data is imported and all of the names are 
converted to lower case.
```{r}
env.df <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/historicalenvdata_final.csv") %>% 
  data.table::fread(data.table = FALSE) %>% 
  dplyr::rename_all(tolower)
```

### Disturbance  (HELP)

A disturbance gradient was created with specific conductance (`conduct`), 
the percentage of forested land in the watershed delineated above the sampling
site (`pct_forest`), and the percentage of developed land in the watershed 
delineated above the sampling site (`pct_developed`). If `conduct` was less than 
or equal to 250 microsiemens __and__ `pct_forest` was greater than or equal to 75%, 
then the site was classified as "reference". If `conduct` was greater than or 
equal to 800 microsiemens __or__ `pct_developed` was greater than or equal to 
25%, then the site was classified as "degraded". Any sites that did not meet these 
specifications were classified as "moderately degraded".

I think these specification need to be updated but I want to get some consensus from the group.
```{r}
env.df <- env.df %>% 
  mutate(gradient = case_when(
    conduct <= 250 & pct_forest >= 75  ~ "reference",
    conduct >= 800 | pct_developed >= 25 ~ "degraded",
    TRUE ~ "moderately degraded"
  ))
table(env.df$gradient)
```

### Join Disturbance Gradient with Taxonomic Data

The `env.df` was subset to only include columns `site`, 
`gradient`, `orig_ibiclassdrainage`, and `ibi_class`. 
`orig_ibiclassdrainage` was renamed to `drainage_size` and 
`ibi_class` was renamed to `basin`.

The subset environmental data, `env.sub`, is joined with the taxonomic 
count data by `site`.
```{r}
env.sub <- env.df %>% 
  dplyr::select(site, gradient, orig_ibiclassdrainage, ibi_class) %>% 
  dplyr::rename(drainage_size = ibi_class,
                basin = orig_ibiclassdrainage)

fish.long <- dplyr::left_join(fish.long, env.sub, by = "site")
```


```{r}
gl.habitat <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_2014_GL_habitat.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)

cs.habitat <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer 2014_chem and susq_habitat.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)

ne.habitat <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer2014_northern and easter_habitat.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)
```

```{r}
new.habitat <- bind_rows(gl.habitat, cs.habitat, ne.habitat) %>% 
  select(-average) %>% 
  gather(site, habitat_value, -Question)
```

```{r}
import_field_data <- function(file.dir) {
  import.df <- file.dir %>% 
    readxl::read_xlsx(sheet = 1, col_names = FALSE) %>% 
  rownames_to_column() %>% 
  gather(var, value, -rowname) %>% 
   spread(rowname, value) %>% 
  select(-var)
  
  names(import.df) <- import.df[1, ]
  final.df <- import.df[-1, ]
  
  return(final.df)
}
```

```{r}
gl.field <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_2014_GL_field data.xlsx") %>% 
  import_field_data()


cs.field <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer 2014_chem and susq_field data sheet.xlsx") %>% 
  import_field_data()

ne.field <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer2014_northern and eastern_field data sheet.xlsx") %>% 
  import_field_data()
```

```{r}
new.field <- bind_rows(gl.field, cs.field, ne.field) 
```

