---
title: "R Notebook"
output: html_notebook
---

## Taxonomic Data
### Historical Dataset
The taxonomic count data is imported below. All of the column names 
are converted to lower case to standardize naming scheme and to make it
easier to type the column names. 

Some of the species names in column `species` 
contained abbreviations or were spelled wrong. The `species` column will be used
to join taxonomic hierarchy information from the Integrated Taxonomic 
Information System (ITIS). Therefore, the names in the `species` columns must be
equivalent to the names found in ITIS. `dplyr` functions `mutate` and `case_when`
where used to update the `species` column. `mutate` allows the 
user to create or overwrite existing columns. `case_when` is equivalent to 
`ifelse` but the syntax is nicer. An example from within the `case_when` call 
below, `species == "n. hog sucker" ~ "northern hog sucker",` would be interpreted 
as replace "n. hog sucker" with "northern hog sucker" in column `species`. The final
line within `case_when`, `TRUE ~ species`, says if the row does not meet any of the 
requirements above then keep the `species` name originally given.

`mutate` and `case_when` are used on columns `water_type`, `tolerance`, 
`native`, and `trophic_class` to make them more descriptive. These 
descriptions will be implemented into the column names during metric 
calculation; therefore, it is beneficial to be more descriptive because 
it is easier to interpret `rich_top_carnivore` rather than `rich_tc`.

Finally, column `species` is re-named to `venacular_name` to be consistent with 
ITIS. `venacular_name` will be used to join the taxonomic counts data frame 
with the taxonomic hierarchy information from ITIS. Additionally, ITIS uses the 
scientific name to fill the `species` column. Joining two data frames that contain 
the same column names but different values within those columns creates issues.

__Question:__ Do we not have any life stage codes (e.g., YOY or fingerlings) for the historical data?

```{r}
#------------------------------------------------------------------------------
# Import fish data.
historical.data <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/all_fish_data.csv") %>% 
  data.table::fread(stringsAsFactors = FALSE) %>% 
  # Change column names to uppercase.
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(
    species = stringr::str_to_lower(species),
    life_stage = case_when(stringr::str_detect(species, "yoy") ~ "yoy",
                           stringr::str_detect(species, "fingerling") ~ "fingerling",
                           TRUE ~ "adult"),
    species = case_when(
      species == "n. hog sucker" ~ "northern hog sucker",
      species == "horneyhead chub" ~ "hornyhead chub",
      species == "mumichog" ~ "mummichog",
      species == "n. redbelly dace" ~ "northern redbelly dace",
      species == "silver redshorse" ~ "silver redhorse",
      species == "e. mudminnow" ~ "eastern mudminnow",
      species == "e. silvery minnow" ~ "eastern silvery minnow",
      species == "e. sand darter" ~ "eastern sand darter",
      species == "trout perch" ~ "trout-perch",
      TRUE ~ species
    ),
    water_type = case_when(
      water_type == "C" ~ "cold_water",
      water_type == "F" ~ "facultative_water",
      water_type == "W" ~ "warm_water",
      TRUE ~ "ERROR"
    ),
    tolerance = case_when(
      tolerance == "I" ~ "intolerant",
      tolerance == "M" ~ "moderately_tolerant",
      tolerance == "T" ~ "tolerant",
      TRUE ~ "ERROR"
    ),
    native = case_when(
      native == "YES" ~ "native",
      native == "NO" ~ "non_native",
      TRUE ~ "ERROR"
    ),
    trophic_class = case_when(
      trophic_class == "BH" ~ "benthic_herbivore",
      trophic_class == "BI" ~ "benthic_insectivore",
      trophic_class == "GF" ~ "generalist",
      trophic_class == "NF" ~ "NF", # What does this stand for?
      trophic_class == "PF" ~ "PF", # What does this stand for?
      trophic_class == "PI" ~ "PI", # What does this stand for?
      trophic_class == "TC" ~ "top_carnivore",
      trophic_class == "WC" ~ "water_column_insectivore",
      TRUE ~ "ERROR"
    )) %>% 
  dplyr::rename(vernacular_name = species)
```

### Transform to Long Format

The data is converted to a long data format using the `tidyr` function 
`gather`. In the script below columns `btnt` through `algy_3`, which are 
site names, are transformed into two columns `site` and `count`. The column 
names are used to populate `site`, while the column values are used to 
populate `count`. Values equal to zero are not useful, and thus, are removed. 
The `site` column character strings are converted to all caps to be 
consistent with the NYSDEC naming scheme. Finally, columns `family`, 
`cold_water`, `warm_water`, `cold_bogs`, and `warm_bogs` are removed from 
the data frame. The `select` function allows you to specify columns you 
want to remove by adding a minus sign in front of the column name.

```{r}
historical.long <- historical.data %>% 
  tidyr::gather(site, count, btnt:algy_3) %>% 
  dplyr::filter(count > 0) %>% 
  dplyr::mutate(site = toupper(site)) %>% 
  dplyr::select(-family, -cold_water, -warm_water,
                -cold_bogs, -warm_bogs)
  
```

### Extract Taxonomic Attributes

```{r}
attr.df <- historical.long %>% 
  select(-site, -count, -life_stage) %>% 
  distinct()

historical.long <- historical.long %>% 
  select(site, vernacular_name, life_stage, count)
```


### 2014 Dataset

```{r}
library(readxl)
gl.df <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_2014_GL_fish.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)

susq.df <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer 2014_chem and susq_fish data.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)

ne.df <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/2014/IBI_summer2014_northeast_fish data.xlsx") %>% 
  readxl::read_xlsx(sheet = 1)
```


```{r}
new.long <- dplyr::full_join(gl.df, susq.df, by = "Site ID") %>% 
  dplyr::full_join(ne.df, by = "Site ID") %>% 
  tidyr::gather(site, count, 2:ncol(.)) %>% 
  dplyr::filter(!is.na(count), count != 0) %>% 
  dplyr::rename(vernacular_name = 'Site ID') %>% 
  dplyr::mutate(
    count = as.numeric(count),
    vernacular_name = stringr::str_to_lower(vernacular_name),
    life_stage = case_when(stringr::str_detect(vernacular_name, "yoy") ~ "yoy",
                           stringr::str_detect(vernacular_name, "fingerling") ~ "fingerling",
                           TRUE ~ "adult"),
    vernacular_name = case_when(
      vernacular_name %in% "n. hog sucker" ~ "northern hog sucker",
      vernacular_name %in% "horneyhead chub" ~ "hornyhead chub", 
      vernacular_name %in% "sculpin" ~ "freshwater sculpins", 
      vernacular_name %in% "n. redbelly dace" ~ "northern redbelly dace",
      vernacular_name %in% "e. brook trout" ~ "brook trout",
      vernacular_name %in% "largemouth bass yoy" ~ "brook trout",
      vernacular_name %in% "brown trout yoy"  ~ "brook trout",
      vernacular_name %in% "brook trout yoy" ~ "brook trout",
      vernacular_name %in% "smallmouth bass yoy" ~ "brook trout",
      vernacular_name %in% "rainbow trout yoy" ~ "brook trout",
      vernacular_name %in% "brown trout fingerlings" ~ "brook trout",
      vernacular_name %in% "sunfish spp." ~ "common sunfishes",
      vernacular_name %in% "unknown shiner"  ~ "minnows",
      vernacular_name %in% "unknown species" ~ "unknown",
      TRUE ~ vernacular_name
    )) %>% 
  filter(!vernacular_name %in% c("minnows", "unknown", "snapping turtles",
                                 "common sunfishes", "freshwater sculpins"))
```

### Join Historical and 2014 Data

```{r}
fish.long <- dplyr::bind_rows(historical.long, new.long)
```

### Join Taxonomic Counts with Taxonomic Attributes

```{r}
fish.long <- dplyr::left_join(fish.long, attr.df, by = "vernacular_name")
```


### Assign Taxonomic Hierarchy

The ITIS database is stored on my computer and accessed using a separate 
script. The vernacular table is imported from the ITIS database and only 
the English vernacular names are kept. A number of Taxonomic Serial Numbers
(TSNs) represented the same vernacular name. This created duplicate entries. 
The vector of TSNs below were removed from the data frame.
```{r, message=FALSE}
source(file.path(rprojroot::find_rstudio_root_file(),
                             "postgres/access_itis.R"))

vern.df <- tbl(conn, "vernaculars") %>% 
  data.frame() %>% 
  mutate(tsn = as.character(tsn),
         vernacular_name = tolower(vernacular_name)) %>%
  filter(language == "English") %>% 
  select(tsn, vernacular_name) %>% 
  filter(!tsn %in% c("163804", "163895", "640068", "163837", "168159", 
                     "164042", "163874", "164044", "163767", "163592", 
                     "164040","650164", "159708", "163938", "163439",
                     "163833", "161100", "161101", "161102", "161103",
                     "161797", "161807", "168506", "168508", "168389",
                     "162141", "162142"))
```

The taxonomic count data frame is joined with the ITIS vernacular data frame 
by the `vernacular_name`. This step assigns each taxon the appropriate TSN. 
The `dplyr` function `left_join` allows the user to specify a column that contains 
shared values between two data frames, and thus, could be used to merge the data 
frames. In the script below, the "left" within `left_join` specifies that only 
`vernacular_names` within `fish.long` should be retained. Any additional 
`vernacular_names` found in `vern.df` will not be included in the new joined 
data frame.
```{r}
fish.long <- dplyr::left_join(fish.long, vern.df, by = "vernacular_name")

dups.df <- fish.long %>% 
  dplyr::group_by(site, vernacular_name, life_stage, count) %>% 
  dplyr::summarise(total = n()) %>% 
  dplyr::arrange(desc(total)) %>% 
  dplyr::filter(total > 1)

if (nrow(dups.df) > 0) DT::datatable(dups.df)

```

The ITIS taxonomic hierarchy data is imported below. Column names 
are converted to lower case and only a subset of columns were retained.
```{r}
# Import fish hierarchy.
hier.df <- file.path(rprojroot::find_rstudio_root_file(),
                             "data/fish_hierarchy.csv") %>% 
  data.table::fread(stringsAsFactors = FALSE,
                             colClasses = "character",
                             na.strings=c("")) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(tsn, rank_name, class, superorder, order, suborder, superfamily, family, subfamily,
         tribe, genus, species) 
```

The taxonomic count data is joined with the ITIS taxonomic hierarchy data 
by the `tsn` column.
```{r}
fish.long <- dplyr::left_join(fish.long, hier.df, by = "tsn")
```
