---
title: "Fishy"
output: 
  html_notebook:
    toc: true
    toc_float: true
---
# Introduction

This document was created using R-Notebook, which aids in creating 
descriptive, reproducible analyses.

## The tidyverse

The packages `dplyr` and `tidyr` are used throughout this analysis. These packages
are from the `tidyverse`, which uses a slightly different syntax than base R. 
The `tidyverse` syntax is intended to reduce typing and to produce more readable
code. Although the syntax takes some time to learn, I find these packages
to be very helpful for manipulating data. Whenever I use these functions in this
document I have tried specify the package that they come from by adding the prefix
`dplyr::` or `tidyr::`. The prefix is not necessary if you load the package using
`library()` but I thought it might help to clarify which package the function
is derived from. 

The `tidyverse` syntax is referred to as Non-Standard Evaluation (NSE). NSE 
allows the user to refer to columns by name without specifying the
data frame for each column and without having to quote the column name.
For example, imagine you want to keep all rows from your data frame,
`df`, where `col.x == TRUE`. In base R, you could write `df[df$col.x == TRUE, ]`
to filter your data frame. The `dplyr` function, `filter`, is more descriptive 
of the action you want to perform and only requires the user to specify the
data frame once. To perform the same filter as above with `dplyr`, 
you would write `filter(df, col.x == TRUE)`. Another useful function from 
`dplyr` is `select`. Imagine `df` has 26 columns and you only want to keep
columns `col.x`, `col.y`, and `col.z`. In base R, you could write 
`df[, c("col.x", "col.y", "col.z")]`. To perform the same action with `dlyr` 
you would use `select(df, col.x, col.y, col.z)`. The `dplyr` function are 
more descriptive and generally require less typing.

Another useful feature of the `tidyverse` is the `magrittr` pipe function, `%>%`. 
The pipe allows the user to chain multiple functions together. The example provided
by Wickham and Grolemund (2016) is `x %>% f(y)` is equivalent to `f(x, y)`. 
Wickham and Grolemund (2016) also suggest reading `%>%` as "then". The following
example would be read as specify `df` as your data frame, then select columns 
`col.x`, `col.y`, and `col.z`, and then filter rows so that `col.x == TRUE`.

```{r eval=FALSE}
df %>% 
select(col.x, col.y, col.z) %>% 
filter(col.x == TRUE)
```

# Preparation

## Import Packages
Load `dplyr` and `tidyr` for data manipulation.
```{r, message=FALSE}
library(dplyr)
library(tidyr)
```

The `Benthos` package is the package I am currently developing. 
`Benthos` is stored on github and can be installed by using 
`devtools::install_github("zsmith27/Benthos", ref = "nse")`. 
If you receive an error when attempting to install from GitHub, 
try running `install.packages(c("devtools","curl", "httr"))` and
then `devtools::install_github("zsmith27/Benthos", ref = "nse")` again. 
If the issue persists please inform me on my GitHub page.

```{r}
library(Benthos)
```

## Taxonomic Data
### Historical Dataset
The taxonomic count data is imported below. All of the column names 
are converted to lower case to standardize naming scheme and to make it
easier to type the column names. 

Some of the species names in column `species` 
contained abbreviations or were spelled wrong. The `species` column will be used
to join taxonomic hierarchy information from the Integrated Taxonomic 
Information System (ITIS). Therefore, the names in the `species` columns must be
equivalent to the names found in ITIS. `dplyr` functions `mutate` and `case_when`
where used to update the `species` column. `mutate` allows the 
user to create or overwrite existing columns. `case_when` is equivalent to 
`ifelse` but the syntax is nicer. An example from within the `case_when` call 
below, `species == "n. hog sucker" ~ "northern hog sucker",` would be interpreted 
as replace "n. hog sucker" with "northern hog sucker" in column `species`. The final
line within `case_when`, `TRUE ~ species`, says if the row does not meet any of the 
requirements above then keep the `species` name originally given.

`mutate` and `case_when` are used on columns `water_type`, `tolerance`, 
`native`, and `trophic_class` to make them more descriptive. These 
descriptions will be implemented into the column names during metric 
calculation; therefore, it is beneficial to be more descriptive because 
it is easier to interpret `rich_top_carnivore` rather than `rich_tc`.

Finally, column `species` is re-named to `venacular_name` to be consistent with 
ITIS. `venacular_name` will be used to join the taxonomic counts data frame 
with the taxonomic hierarchy information from ITIS. Additionally, ITIS uses the 
scientific name to fill the `species` column. Joining two data frames that contain 
the same column names but different values within those columns creates issues.

__Question:__ Do we not have any life stage codes (e.g., YOY or fingerlings) for the historical data?

```{r}
#------------------------------------------------------------------------------
# Import fish data.
historical.data <- read.csv("data/all_fish_data.csv", stringsAsFactors = FALSE) %>% 
  # Change column names to uppercase.
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(
    species = stringr::str_to_lower(species),
    life_stage = case_when(stringr::str_detect(species, "yoy") ~ "yoy",
                           stringr::str_detect(species, "fingerling") ~ "fingerling",
                           TRUE ~ "adult"),
    species = case_when(
      species == "n. hog sucker" ~ "northern hog sucker",
      species == "horneyhead chub" ~ "hornyhead chub",
      species == "mumichog" ~ "mummichog",
      species == "n. redbelly dace" ~ "northern redbelly dace",
      species == "silver redshorse" ~ "silver redhorse",
      species == "e. mudminnow" ~ "eastern mudminnow",
      species == "e. silvery minnow" ~ "eastern silvery minnow",
      species == "e. sand darter" ~ "eastern sand darter",
      species == "trout perch" ~ "trout-perch",
      TRUE ~ species
    ),
    water_type = case_when(
      water_type == "C" ~ "cold_water",
      water_type == "F" ~ "facultative_water",
      water_type == "W" ~ "warm_water",
      TRUE ~ "ERROR"
    ),
    tolerance = case_when(
      tolerance == "I" ~ "intolerant",
      tolerance == "M" ~ "moderately_tolerant",
      tolerance == "T" ~ "tolerant",
      TRUE ~ "ERROR"
    ),
    native = case_when(
      native == "YES" ~ "native",
      native == "NO" ~ "non_native",
      TRUE ~ "ERROR"
    ),
    trophic_class = case_when(
      trophic_class == "BH" ~ "benthic_herbivore",
      trophic_class == "BI" ~ "benthic_insectivore",
      trophic_class == "GF" ~ "generalist",
      trophic_class == "NF" ~ "NF", # What does this stand for?
      trophic_class == "PF" ~ "PF", # What does this stand for?
      trophic_class == "PI" ~ "PI", # What does this stand for?
      trophic_class == "TC" ~ "top_carnivore",
      trophic_class == "WC" ~ "water_column_insectivore",
      TRUE ~ "ERROR"
    )) %>% 
  dplyr::rename(vernacular_name = species)
```

### Transform to Long Format

The data is converted to a long data format using the `tidyr` function 
`gather`. In the script below columns `btnt` through `algy_3`, which are 
site names, are transformed into two columns `site` and `count`. The column 
names are used to populate `site`, while the column values are used to 
populate `count`. Values equal to zero are not useful, and thus, are removed. 
The `site` column character strings are converted to all caps to be 
consistent with the NYSDEC naming scheme. Finally, columns `family`, 
`cold_water`, `warm_water`, `cold_bogs`, and `warm_bogs` are removed from 
the data frame. The `select` function allows you to specify columns you 
want to remove by adding a minus sign in front of the column name.

```{r}
historical.long <- historical.data %>% 
  tidyr::gather(site, count, btnt:algy_3) %>% 
  dplyr::filter(count > 0) %>% 
  dplyr::mutate(site = toupper(site)) %>% 
  dplyr::select(-family, -cold_water, -warm_water,
                -cold_bogs, -warm_bogs)
  
```

### Extract Taxonomic Attributes

```{r}
attr.df <- historical.long %>% 
  select(-site, -count, -life_stage) %>% 
  distinct()

historical.long <- historical.long %>% 
  select(site, vernacular_name, life_stage, count)
```


### 2014 Dataset

```{r}
library(readxl)
gl.df <- readxl::read_xlsx("data/2014/IBI_2014_GL_fish.xlsx", sheet = 1)
susq.df <- readxl::read_xlsx("data/2014/IBI_summer 2014_chem and susq_fish data.xlsx", sheet = 1)
ne.df <- readxl::read_xlsx("data/2014/IBI_summer2014_northeast_fish data.xlsx", sheet = 1)
```


```{r}
new.long <- dplyr::full_join(gl.df, susq.df, by = "Site ID") %>% 
  dplyr::full_join(ne.df, by = "Site ID") %>% 
  tidyr::gather(site, count, 2:ncol(.)) %>% 
  dplyr::filter(!is.na(count), count != 0) %>% 
  dplyr::rename(vernacular_name = 'Site ID') %>% 
  dplyr::mutate(
    count = as.numeric(count),
    vernacular_name = stringr::str_to_lower(vernacular_name),
    life_stage = case_when(stringr::str_detect(vernacular_name, "yoy") ~ "yoy",
                           stringr::str_detect(vernacular_name, "fingerling") ~ "fingerling",
                           TRUE ~ "adult"),
    vernacular_name = case_when(
      vernacular_name %in% "n. hog sucker" ~ "northern hog sucker",
      vernacular_name %in% "horneyhead chub" ~ "hornyhead chub", 
      vernacular_name %in% "sculpin" ~ "freshwater sculpins", 
      vernacular_name %in% "n. redbelly dace" ~ "northern redbelly dace",
      vernacular_name %in% "e. brook trout" ~ "brook trout",
      vernacular_name %in% "largemouth bass yoy" ~ "brook trout",
      vernacular_name %in% "brown trout yoy"  ~ "brook trout",
      vernacular_name %in% "brook trout yoy" ~ "brook trout",
      vernacular_name %in% "smallmouth bass yoy" ~ "brook trout",
      vernacular_name %in% "rainbow trout yoy" ~ "brook trout",
      vernacular_name %in% "brown trout fingerlings" ~ "brook trout",
      vernacular_name %in% "sunfish spp." ~ "common sunfishes",
      vernacular_name %in% "unknown shiner"  ~ "minnows",
      vernacular_name %in% "unknown species" ~ "unknown",
      TRUE ~ vernacular_name
    )) %>% 
  filter(!vernacular_name %in% c("minnows", "unknown", "snapping turtles",
                                 "common sunfishes", "freshwater sculpins"))
```

### Join Historical and 2014 Data

```{r}
fish.long <- dplyr::bind_rows(historical.long, new.long)
```

### Join Taxonomic Counts with Taxonomic Attributes

```{r}
fish.long <- dplyr::left_join(fish.long, attr.df, by = "vernacular_name")
```


### Assign Taxonomic Hierarchy

The ITIS database is stored on my computer and accessed using a separate 
script. The vernacular table is imported from the ITIS database and only 
the English vernacular names are kept. A number of Taxonomic Serial Numbers
(TSNs) represented the same vernacular name. This created duplicate entries. 
The vector of TSNs below were removed from the data frame.
```{r, message=FALSE}
source("postgres/access_itis.R")

vern.df <- tbl(conn, "vernaculars") %>% 
  data.frame() %>% 
  mutate(tsn = as.character(tsn),
         vernacular_name = tolower(vernacular_name)) %>%
  filter(language == "English") %>% 
  select(tsn, vernacular_name) %>% 
  filter(!tsn %in% c("163804", "163895", "640068", "163837", "168159", 
                     "164042", "163874", "164044", "163767", "163592", 
                     "164040","650164", "159708", "163938", "163439",
                     "163833", "161100", "161101", "161102", "161103",
                     "161797", "161807", "168506", "168508", "168389",
                     "162141", "162142"))
```

The taxonomic count data frame is joined with the ITIS vernacular data frame 
by the `vernacular_name`. This step assigns each taxon the appropriate TSN. 
The `dplyr` function `left_join` allows the user to specify a column that contains 
shared values between two data frames, and thus, could be used to merge the data 
frames. In the script below, the "left" within `left_join` specifies that only 
`vernacular_names` within `fish.long` should be retained. Any additional 
`vernacular_names` found in `vern.df` will not be included in the new joined 
data frame.
```{r}
fish.long <- dplyr::left_join(fish.long, vern.df, by = "vernacular_name")

dups.df <- fish.long %>% 
  dplyr::group_by(site, vernacular_name, life_stage, count) %>% 
  dplyr::summarise(total = n()) %>% 
  dplyr::arrange(desc(total)) %>% 
  dplyr::filter(total > 1)

if (nrow(dups.df) > 0) DT::datatable(dups.df)

```

The ITIS taxonomic hierarchy data is imported below. Column names 
are converted to lower case and only a subset of columns were retained.
```{r}
# Import fish hierarchy.
hier.df <- data.table::fread("data/fish_hierarchy.csv",
                             stringsAsFactors = FALSE,
                             colClasses = "character",
                             na.strings=c("")) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(tsn, rank_name, class, superorder, order, suborder, superfamily, family, subfamily,
         tribe, genus, species) 
```

The taxonomic count data is joined with the ITIS taxonomic hierarchy data 
by the `tsn` column.
```{r}
fish.long <- dplyr::left_join(fish.long, hier.df, by = "tsn")
```

## Station Data

### Import

The station and environmental data is imported and all of the names are 
converted to lower case.
```{r}
env.df <- data.table::fread("data/historicalenvdata_final.csv",
                              data.table = FALSE) %>% 
  dplyr::rename_all(tolower)
```

### Disturbance  (HELP)

A disturbance gradient was created with specific conductance (`conduct`), 
the percentage of forested land in the watershed delineated above the sampling
site (`pct_forest`), and the percentage of developed land in the watershed 
delineated above the sampling site (`pct_developed`). If `conduct` was less than 
or equal to 250 microsiemens __and__ `pct_forest` was greater than or equal to 75%, 
then the site was classified as "reference". If `conduct` was greater than or 
equal to 800 microsiemens __or__ `pct_developed` was greater than or equal to 
25%, then the site was classified as "degraded". Any sites that did not meet these 
specifications were classified as "moderately degraded".

I think these specification need to be updated but I want to get some consensus from the group.
```{r}
env.df <- env.df %>% 
  mutate(gradient = case_when(
    conduct <= 250 & pct_forest >= 75  ~ "reference",
    conduct >= 800 | pct_developed >= 25 ~ "degraded",
    TRUE ~ "moderately degraded"
  ))
table(env.df$gradient)
```

### Join Disturbance Gradient with Taxonomic Data

The `env.df` was subset to only include columns `site`, 
`gradient`, `orig_ibiclassdrainage`, and `ibi_class`. 
`orig_ibiclassdrainage` was renamed to `drainage_size` and 
`ibi_class` was renamed to `basin`.

The subset environmental data, `env.sub`, is joined with the taxonomic 
count data by `site`.
```{r}
env.sub <- env.df %>% 
  dplyr::select(site, gradient, orig_ibiclassdrainage, ibi_class) %>% 
  dplyr::rename(drainage_size = ibi_class,
                basin = orig_ibiclassdrainage)

fish.long <- dplyr::left_join(fish.long, env.sub, by = "site")
```


# Metric Calculation

Metrics are calculated using the `Benthos` package that I am developing.

## Sequence Metrics

The `Benthos` function `seq_taxa` allows the user to sequence across a 
specified vector of column names and calculate the percentage, abundance, 
or richness of each taxon in the specified columns. The vector of column 
names currently being used is `hier.vec`. You will notice that this vector 
does not just include taxonomic columns, but also, taxonomic attribute columns. 
Each taxonomic attribute is treated like a unique taxon, and therefore, the 
percentage, abundance, and richness of each taxonomic attribute can also be 
calculated. The `seq_taxa` function can only do one `job` at a time. The `dplyr` 
function `bind_cols` is used to append columns together from each `seq_taxa` call. 
Each metric calculated with the `seq_taxa` function will be named with a prefix 
designated to the specified `job` ("pct" = percentage, "abund" = abundance, and 
"rich" = richness) followed by the taxonomic name or taxonomic attribute name.

```{r, message=FALSE}
hier.vec <- c("water_type", "tolerance", "native", "trophic_class",
              "class", "superorder", "order", "suborder", "superfamily",
              "family", "subfamily", "tribe", "genus", "species")

seq.df <- Benthos::seq_taxa(fish.long, site, count, hier.vec, job = "pct") %>% 
  dplyr::bind_cols(Benthos::seq_taxa(fish.long, site, count, hier.vec, job = "abund")) %>% 
  dplyr::bind_cols(Benthos::seq_taxa(fish.long, site, count, hier.vec[!hier.vec %in% "species"],
                     high.taxa.col = "species", job = "rich"))
```

## Additional Metrics

Create a new data frame to store metric values. Keep only the columns 
that will be useful in subsequent analyses.
```{r}
metrics.df <- fish.long %>% 
  select(site, gradient, basin, drainage_size) %>% 
  distinct()
```

The `seq_taxa` function cannot be used to calculate all metrics. Below 
total species richness, weighted species richness, total abundance, and 
the percent of dominant taxa (1, 2, 3, 4, and 5). `pct_`, `abund_`, and 
`rich_` `insectivores` was calculated as the sum of the corresponding 
`benthic_insectivore`, `water_column_insectivore`, and `pi` metrics calculated 
with `seq_taxa`.
```{r}
#==============================================================================
metrics.df <- dplyr::bind_cols(metrics.df, seq.df)

metrics.df <- metrics.df %>% 
  mutate(richness = taxa_rich(fish.long, site, count, family, species)) %>% 
  left_join(env.df[, c("site", "width")], by = "site") %>% 
#  mutate(weighted_richness =
#           case_when(
#             width <= 4 ~ richness + 2,
#             width > 4 & width <= 9 ~ as.double(richness),
#             width > 9 & width <= 19 ~ richness - 2,
#             width > 19 ~ richness - 4,
#             is.na(width) ~ as.numeric(NA),
#             TRUE ~ as.numeric(NA)
#           )) %>% 
  select(-width) %>% 
  mutate(abundance = taxa_abund(fish.long, site, count, species),
         shannon = taxa_div(fish.long, site, count, species, job = "shannon", base = 2),
         simpson = taxa_div(fish.long, site, count, species, job = "simpson"),
         hill_2 = taxa_div(fish.long, site, count, species, job = "hill", q = 2),
         pct_dom1 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 1),
         pct_dom2 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 2),
         pct_dom3 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 3),
         pct_dom4 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 4),
         pct_dom5 = pct_dom(fish.long,
                        unique.id.col = site,
                        count.col = count,
                        taxa.col = species,
                        dom.level = 5),
         pct_insectivores = sum(pct_benthic_insectivore,
                                pct_water_column_insectivore,
                                pct_pi),
         abund_insectivores = sum(abund_benthic_insectivore,
                                abund_water_column_insectivore,
                                abund_pi),
         rich_insectivores = sum(rich_benthic_insectivore,
                                rich_water_column_insectivore,
                                rich_pi))
```


`r ncol(metrics.df) - 4` metrics were calculated.

# Metric Sensitivity

To calculate metric sensitivity we will install the other package I 
am developing, `mmi`. This package is also available on GitHub and can be 
installed with `devtools::install_github("zsmith27/MMI", ref = "build")`.
If you have difficulties installing the package please try running 
`install.packages(c("devtools","curl", "httr"))` and then run 
`devtools::install_github("zsmith27/MMI", ref = "build")` again. If 
the issue persists, please contact me through my GitHub page.
```{r, message=FALSE}
library(MMI)
```

## State-Wide Index

Metric sensitivity was calculated without splitting the data by basin or drainage area. 
This spatial resolution would create a single index for New York State. 
```{r}
#table.df <- table(metrics.df$gradient) %>% 
#  data.frame() %>% 
  
```

```{r}
sens.all <- MMI::sensitivity(metrics.df,
                        first.metric = "pct_warm_water",
                        condition.colname = "gradient",
                        ref.cond = "reference",
                        deg.cond = "degraded",
                        method = "ALL") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::arrange(desc(de_sensitivity))
```

### State-Wide Metric Sensitivity

The table shows a portion of the output from `MMI` function `sensitivity`. 

* `disturbance` =
    + "INCREASE" if the metric increases with disturbance
    + "DECREASE" if the metric decreases with disturbance
    + "EQUAL" if there is no change due to disturbance
* `barbour_sensitivity` = the barbour visual method was automated.
    + 0 = Overlap of the reference median with the degraded IQR and overlap of the degraded median with the reference IQR
    + 1 = Overlap of either one distributions median with the other distributions IQR
    + 2 = Overlap of the reference IQR with the degraded IQR but no overlap of median values
    + 3 = No overlap of the IQR of reference and degraded distributions
* `de_sensitivity` = Discrimination Efficiency (DE) is the most popular method for calculating metric sensitivity. For metrics that decrease with disturbance, DE is the percentage of degraded samples correctly identified below the reference 25th percentile. For metrics that increase with disturbance, DE is the percentage of degraded samples correctly identified above the reference 75th percentile. 100 is the best possible value and 0 is the worst. Typically, values greater than or equal to 75 are considered good.
* `bde_sensitivity` = Balanced Discrimination Efficiency (BDE) is a measure that I have been working on with my boss Claire Buchannan. The standard DE method is kind of arbitrary. Our method tests every reference percentile as a potential threshold for splitting reference and degraded distributions. DE only tests the reference 25th or 75th percentile as a splitting point.  We also aim to balance the number of reference correctly identified with the threshold with the number of degraded correctly identified.  I think this method is great but it is still in development; therefore, it is probably safer to just stick with Barbour et al.'s (1996) method and the DE method.

Each table represents only the top 10 performing metrics.
```{r}
sens.all %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

## Basin Indices

This spatial resolution divides NYS into four basins and is consistent 
with Anne's thesis.
```{r}
table.df <- table(metrics.df$basin, metrics.df$gradient) %>% 
  data.frame() %>% 
  dplyr::rename(basin = Var1,
                gradient = Var2) %>% 
  tidyr::spread(gradient, Freq) %>% 
  dplyr::select(basin, reference, 'moderately degraded', degraded)
DT::datatable(table.df, options = list(dom = 't'))
```


```{r}
basin.list <- lapply(unique(metrics.df$basin), function(basin.i) {
  sub.df <- dplyr::filter(metrics.df, basin == basin.i)
  sens.df <- MMI::sensitivity(sub.df,
                        first.metric = "pct_warm_water",
                        condition.colname = "gradient",
                        ref.cond = "reference",
                        deg.cond = "degraded",
                        method = "ALL") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::mutate(basin = basin.i) %>% 
  dplyr::select(basin, everything())
})

sens.basin <- dplyr::bind_rows(basin.list)
```

### North/East Metric Sensitivity
```{r}
sens.basin %>% 
  dplyr::filter(basin == "North/East") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

### Great Lakes Metric Sensitivity
```{r}
sens.basin %>% 
  dplyr::filter(basin == "Great Lakes") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

### Chesapeake Metric Sensitivity
```{r}
sens.basin %>% 
  dplyr::filter(basin == "Chesapeake") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

### Mississippi Metric Sensitivity
```{r}
sens.basin %>% 
  dplyr::filter(basin == "Mississippi") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```


## Drainage Size Indices

This spatial resolution divides NYS up into two drainage sizes 
(i.e., "Headwater" and "Mid-Order") and the "Mississippi basin". 
This spatial resolution is based on the new groups specified by A.J.
```{r}
table.df <- table(metrics.df$drainage_size, metrics.df$gradient) %>% 
  data.frame() %>% 
  dplyr::rename(drainage_size = Var1,
                gradient = Var2) %>% 
  tidyr::spread(gradient, Freq) %>% 
  dplyr::select(drainage_size, reference, 'moderately degraded', degraded)
DT::datatable(table.df, options = list(dom = 't'))
```


```{r}
drainage.list <- lapply(unique(metrics.df$drainage_size), function(drainage.i) {
  sub.df <- dplyr::filter(metrics.df, drainage_size == drainage.i)
  sens.df <- MMI::sensitivity(sub.df,
                        first.metric = "pct_warm_water",
                        condition.colname = "gradient",
                        ref.cond = "reference",
                        deg.cond = "degraded",
                        method = "ALL") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::mutate(drainage_size = drainage.i) %>% 
  dplyr::select(drainage_size, everything())
})

sens.drainage <- dplyr::bind_rows(drainage.list)

```

### Headwater Metric Sensitivity
```{r}
sens.drainage %>% 
  dplyr::filter(drainage_size == "Mississippi") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

### Mid-Order Metric Sensitivity
```{r}
sens.drainage %>% 
  dplyr::filter(drainage_size == "Mid-Order") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

### Mississippi Metric Sensitivity
```{r}
sens.drainage %>% 
  dplyr::filter(drainage_size == "Headwater") %>% 
  dplyr::arrange(desc(de_sensitivity)) %>% 
  dplyr::select(1:6) %>% 
  dplyr::slice(1:10) %>% 
  DT::datatable(options = list(dom = 't'))
```

